---
title: "ML_Project4"
author: "Breanna Barton"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
library('readxl')
library(ROSE)
library(car)
library(rsample)
library(randomForest)
library(ranger)  
library(caret) 
library(ROCR)
library(rpart.plot)
library(rpart)
library(ipred)
library(dplyr)

set.seed(171)
setwd("C:/Users/Peter Shoemaker/OneDrive/Desktop/Grad School/Machine Learning I/Project 4")
# load data
emp <- read_excel("Employee_Data_Project.xlsx")
head(emp)

# training and testing data
trainrows <- sample(rownames(emp), dim(emp)[1]*0.7)
train <- emp[trainrows, ]
testrows <- setdiff(rownames(emp), trainrows)
test <- emp[testrows, ]

# data cleaning
train$JobSatisfaction <- as.numeric(train$JobSatisfaction)
train$EnvironmentSatisfaction <- as.numeric(train$EnvironmentSatisfaction)
train$TotalWorkingYears <- as.numeric(train$TotalWorkingYears)
train$NumCompaniesWorked <- as.numeric(train$NumCompaniesWorked)



test$JobSatisfaction <- as.numeric(test$JobSatisfaction)
test$EnvironmentSatisfaction <- as.numeric(test$EnvironmentSatisfaction)
test$TotalWorkingYears <- as.numeric(test$TotalWorkingYears)
test$NumCompaniesWorked <- as.numeric(test$NumCompaniesWorked)

#modifications
train$BusinessTravel <- as.factor(train$BusinessTravel)
test$BusinessTravel <- as.factor(test$BusinessTravel)
train$MaritalStatus <- as.factor(train$MaritalStatus)
test$MaritalStatus <- as.factor(test$MaritalStatus)
train$Gender <- as.factor(train$Gender)
test$Gender <- as.factor(test$Gender)
train$Single <- ifelse(train$MaritalStatus=='Single',1,0)
test$Single <- ifelse(test$MaritalStatus=='Single',1,0)
train$FreqTravel <- ifelse(train$BusinessTravel=='Travel_Frequently',1,0)
test$FreqTravel <- ifelse(test$BusinessTravel=='Travel_Frequently',1,0)


train <- na.omit(train)
test <- na.omit(test)

colSums(is.na(train))
colSums(is.na(test))

# Attrition - change yes to 1 and no to 0
train$Attrition <- ifelse(train$Attrition =="Yes", 1, 0)
test$Attrition <- ifelse(test$Attrition =="Yes", 1, 0)

str(train)
str(test)
```


LOGISTIC REGRESSION: What factors lead to employee attrition within the company? Use Logistic Regression to determine the model fit. What are the top 3 variables based on the logistic regression model that is most important for the management to address right away to curb attrition? (15 points)

```{r}
# Downsampling

data_balanced_under <- ovun.sample(Attrition ~ ., data = train, method = "under",N = 1000)$data
table(data_balanced_under$Attrition)

data_balanced_under$NumCompaniesWorked <- as.numeric(data_balanced_under$NumCompaniesWorked)
data_balanced_under$BusinessTravel <- as.factor(data_balanced_under$BusinessTravel)
data_balanced_under$MaritalStatus <- as.factor(data_balanced_under$MaritalStatus)
data_balanced_under$Gender <- as.factor(data_balanced_under$Gender)
data_balanced_under$Single <- ifelse(data_balanced_under$MaritalStatus=='Single',1,0)
data_balanced_under$FreqTravel <- ifelse(data_balanced_under$BusinessTravel=='Travel_Frequently',1,0)


train = data_balanced_under


# logistic model
Logistic_model10 <- glm(formula = Attrition ~ Age + JobSatisfaction + FreqTravel + Single + NumCompaniesWorked + TrainingTimesLastYear + YearsWithCurrManager + EnvironmentSatisfaction, family = binomial(), data = data_balanced_under)

summary(Logistic_model10)

stepmodel4 = step(Logistic_model10, direction="both")

formula(stepmodel4)
summary(stepmodel4)


test$Single <- ifelse(test$MaritalStatus=='Single',1,0)
test$FreqTravel <- ifelse(test$BusinessTravel=='Travel_Frequently',1,0)

vif(stepmodel4)

pred <- predict(object = stepmodel4, newdata = test, type = "response")

head(pred)
summary(pred)

#Lets set the threshold to 0.50 for predicting into 1
test$predicted<-ifelse(pred>=0.5, 1, 0)
head(test$predicted)
 
# Accuracy of model 
table(test$Attrition, test$predicted)

confusionMatrix(data = as.factor(test$predicted),
                reference =  as.factor(test$Attrition),
                positive = "1")


#ROC
roc_pred <- prediction(predictions = pred  , labels = test$Attrition)
roc_perf <- performance(roc_pred , "tpr" , "fpr")
plot(roc_perf,
     colorize = TRUE,
     print.cutoffs.at= seq(0,1,0.05),
     text.adj=c(-0.2,1.7))

# AUC (TWO WAYS)
logistic_roc = roc.curve(test$Attrition, pred)
auc_logistic_regression = as.numeric(performance(roc_pred, "auc")@y.values)

# RMSE for logistic regression
RMSE(pred = pred, obs = test$Attrition)
```


DECISION TREES: What factors lead to employee attrition within the company? Use Decision Trees to determine the response. What are the top 3 variables based on the Decision Tree model that is most important for the management to address right away to curb attrition? (30 points)

```{r}

#### INITIAL DECISION TREE #####

# Simple Regression Tree
# CART Modeling - Classification and regression trees
# method = "class" for classification tree
# method = "anova" for regression tree
m1 <- rpart(
  formula = Attrition ~ .,
  data    = train,
  method  = "class"
)

m1

varImp(m1)

#Plot the tree
rpart.plot(m1)

#ANOVA - Continuous Response
#shows the predicted value (is this the probability of attrition? or the error?)
#shows percentage of observations in the node

#CLASS - Binary Response
#shows the predicted class
#shows the predicted probability
#shows the percentage of observatiosn in the node 

#Plot cross-validation error with cost complexity or α value
plotcp(m1)

#display cp table
printcp(m1)

#plot approximate R-squared and relative error for different splits (2 plots), labels are only appropriate for the "anova" method
rsq.rpart(m1)

#detailed results including surrogate splits
summary(m1)
```

```{r}
#### TUNING ####

#Instead of manual tuning, we can do a grid search over range of minsplits and maxdepths
hyper_grid <- expand.grid(
  minsplit = seq(5, 10, 1), #5-20, incrementing by 1
  maxdepth = seq(3, 8, 1) #7-12, incrementing by 1
)

head(hyper_grid)

# total number of combinations
nrow(hyper_grid)

#To automate the modeling, we simply set up a for loop and iterate through each minsplit and maxdepth combination 
#We save each model into its own list item

models <- list()

for (i in 1:nrow(hyper_grid)) {
  # get minsplit, maxdepth values at row i
  minsplit <- hyper_grid$minsplit[i]
  maxdepth <- hyper_grid$maxdepth[i]
  
  # train a model and store in the list
  models[[i]] <- rpart(
    formula = Attrition ~ .,
    data    = train,
    method  = "anova",
    control = list(minsplit = minsplit, maxdepth = maxdepth)
  )
}

# function to get optimal cp
get_cp <- function(x) {
  min    <- which.min(x$cptable[, "xerror"])
  cp <- x$cptable[min, "CP"] 
}

# function to get minimum error
get_min_error <- function(x) {
  min    <- which.min(x$cptable[, "xerror"])
  xerror <- x$cptable[min, "xerror"] 
}

set.seed(171)
hyper_grid %>%
  mutate(
    cp    = purrr::map_dbl(models, get_cp),
    error = purrr::map_dbl(models, get_min_error)
  ) %>%
  arrange(error) %>%
  top_n(-5, wt = error)


```


```{r}
### PREDICTION ON TEST SET ###

# we want to choose the lowest error value and use the corresponding minsplit and maxdepth values
# grid search helps obtain the more optimal model

# predict on test set with optimal numbers

optimal_tree <- rpart(
  formula = Attrition ~ .,
  data    = train,
  method  = "anova",
  control = list(minsplit = 5, maxdepth = 6, cp = 0.01)
)

pred_tree <- predict(optimal_tree, newdata = test)

RMSE(pred = pred_tree, obs = test$Attrition)
# with values of this gives the final RMSE of 0.3628671 which suggests that...

optimal_tree

rpart.plot(optimal_tree)

varImp(optimal_tree)


#PETER CODE
pred_tree <-ifelse(pred_tree >= 0.5, 1, 0)
confusionMatrix(data = as.factor(pred_tree),
                reference =  as.factor(test$Attrition),
                positive = "1")
roc_pred2 <- prediction(predictions = pred_tree  , labels = test$Attrition)
# AUC (TWO WAYS)
tree_roc = roc.curve(test$Attrition, pred_tree)
auc_tree = as.numeric(performance(roc_pred2, "auc")@y.values)


```

EXTENDING DECISION TREES: Use Bagging and Random Forest methods to determine employee attrition (30 points)

```{r}

#############BAGGING###############
#make bootstrapping reproducible
set.seed(171)

# train bagged model
bagged_m1 <- bagging(
  formula = Attrition ~ .,
  data    = train,
  coob    = TRUE
)

bagged_m1

# assess 10-50 bagged trees
ntree <- 10:50

# create empty vector to store OOB RMSE values
rmse <- vector(mode = "numeric", length = length(ntree))

for (i in seq_along(ntree)) {
  # reproducibility
  set.seed(171)
  
  # perform bagged model
  model <- bagging(
    formula = Attrition ~ .,
    data    = train,
    coob    = TRUE,
    nbagg   = ntree[i]
  )
  # get OOB error
  rmse[i] <- model$err
}

plot(ntree, rmse, type = 'l', lwd = 2)
abline(v = 25, col = "red", lty = "dashed")



#BAGGING WITH CARET PACKAGE
#Bagging with caret helps to perform cross-validation easily and assess variable importance across bagged trees
#10-fold cross-validation

train$Attrition <- as.factor(train$Attrition)
str(train)


# Specify 10-fold cross validation
ctrl <- trainControl(method = "cv",  number = 10) 

# CV bagged model
bagged_cv <- train(
  Attrition ~ .,
  data = train,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE, 
  na.action=na.omit
)

# assess results
bagged_cv

# plot most important variables
plot(varImp(bagged_cv), 20)  

#Predicting to test set
#pred2 <- predict(bagged_cv, test, na.action=na.roughfix)
pred2 <- as.numeric(predict(bagged_cv, test))
pred2 <-ifelse(pred2 >= 1.5, 1, 0)
RMSE(pred2, test$Attrition)

#PETER CODE
bagging_confusion = confusionMatrix(data = as.factor(pred2),
                reference =  as.factor(test$Attrition),
                positive = "1")
bagging_roc = roc.curve(test$Attrition, pred2)
roc_pred2 <- prediction(predictions = pred2  , labels = test$Attrition)
auc_bagging = as.numeric(performance(roc_pred2, "auc")@y.values)


#############RANDOM FOREST###############

train$Attrition <- as.numeric(train$Attrition)
str(train)

#use set.seed for reproducibility
set.seed(171)

# default RF model
#you can either impute missing values before model or use na.action = na.roughfix for mean/mode values
#of use na.action=na.omit if you would like to omit all columns that have at least one missing value
rf1 <- randomForest(
  formula = Attrition ~ .,
  data    = train
  #,na.action=na.roughfix
)

#na.action=na.roughfix -> imputes missing values using median or mode...only works for numeric or factor 

rf1

#Plotting the model will illustrate the error rate as we average across more trees
plot(rf1)

#Error rate based on the Out of Bag (OOB) sample error can be accessed at m1$mse 
#No. of trees with lowest MSE
which.min(rf1$mse)

# RMSE of this optimal random forest
sqrt(rf1$mse[which.min(rf1$mse)])

summary(rf1)

varImp(rf1)%>%
  top_n(10)%>%
  arrange(desc(Overall))

#PETER CODE
pred3 <- predict(rf1, test)
pred3 = pred3 - 1
pred3 <-ifelse(pred3>=0.5, 1, 0)
rf_confusion = confusionMatrix(data = as.factor(pred3),
                reference =  as.factor(test$Attrition),
                positive = "1")

random_forest_roc = roc.curve(test$Attrition, pred3)
roc_pred3 <- prediction(predictionse = pred3  , labels = test$Attrition)
auc_random_forest = as.numeric(performance(roc_pred3, "auc")@y.values)

```


